---
title: "The lorentz package: special relativity in R"
output:
  github_document:
    pandoc_args: --webtex
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![Build Status](https://travis-ci.org/RobinHankin/lorentz.svg?branch=master)](https://travis-ci.org/RobinHankin/lorentz)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/lorentz)](https://cran.r-project.org/package=lorentz)
[![Rdoc](http://www.rdocumentation.org/badges/version/lorentz)](http://www.rdocumentation.org/packages/lorentz)
<!-- badges: end -->


# Overview


The ```lorentz``` package furnishes some R-centric functionality for
special relativity.  Lorentz transformations of four-vectors are
handled and some functionality for the stress energy tensor is given.
The package deals with four-momentum and has facilities for dealing
with photons and mirrors in relativistic situations.  A detailed
vignette is provided in the package.

The original motivation for the package was the investigation of the
(nonassociative) gyrogroup structure of relativistic three-velocities
under Einsteinian velocity composition.  Natural R idiom may be used
to manipulate vectors of three-velocities, although one must be
careful with brackets.


# Installation

To install the most recent stable version on CRAN, use
```install.packages()``` at the R prompt:

```
R> install.packages("lorentz")
```


To install the current development version use ```devtools```:

```
R> devtools::install_github("RobinHankin/lorentz")
```

And then to load the package use ```library()```:


```{r,message=FALSE}
library("lorentz")
```


# The ```lorentz``` package in use

The package furnishes natural R idiom for working with
three-velocities, four-velocities, and Lorentz transformations as
four-by-four matrices.  Although natural units in which $c=1$ are used
by default, this can be changed.


```{r}
 u <- as.3vel(c(0.6,0,0))  # define a three-velocity, 0.6c to the right
 u

as.4vel(u)    # convert to a four-velocity:
gam(u)  # calculate the gamma term

B <- boost(as.3vel(c(0.6,0,0))) # give the Lorentz transformation
B

 B %*% (1:4)  # Lorentz transform of an arbitrary four-vector
```

The package is fully vectorized:

```{r}
 set.seed(0)
 options(digits=3)
 # generate 5 random three-velocities:
 (u <- r3vel(5))
 # calculate the gamma correction term:
 gam(u)

 # add a velocity of 0.9c in the x-direction:
 v <- as.3vel(c(0.9,0,0))
 v+u


 # convert u to a four-velocity:
 as.4vel(u)

 # use four-velocities to effect the same transformation:
 w <- as.4vel(u) %*% boost(-v)
 as.3vel(w)
```


### Creation of three velocities:

Three-velocites behave in interesting and counter-intuitive ways.


```{r}
 u <- as.3vel(c(0.2,0.4,0.1))   # single three-velocity
 v <- r3vel(4,0.9)              # 4 random three-velocities with speed 0.9
 w <- as.3vel(c(-0.5,0.1,0.3))  # single three-velocity
```

The three-velocity addition law is given by Ungar.

Then we can see that velocity addition is not commutative:

```{r}
 u+v
 v+u
 (u+v)-(v+u)
```

Observe that the difference between ```u+v``` and ```v+u``` is not
"small" in any sense. Commutativity is replaced with gyrocommutatitivity:

```{r}
# Compare two different ways of calculating the same thing:
 (u+v) - gyr(u,v,v+u)  

# The other way round:
 (v+u) - gyr(v,u,u+v)
``` 

(that is, zero to numerical accuracy)

### Nonassociativity

It would be reasonable to expect that ```u+(v+w)==(u+v)+w```.
However, this is not the case:
 
```{r}
 ((u+v)+w) - (u+(v+w))
``` 

(that is, significant departure from associativity).
Associativity is replaced with gyroassociativity:

```{r}
 (u+(v+w)) - ((u+v)+gyr(u,v,w))
 ((u+v)+w) - (u+(v+gyr(v,u,w)))
```

(zero to numerical accuracy).


# References

The most concise reference is

*  A. A. Ungar 2006. _Thomas precession: a kinematic effect of the algebra of Einstein's velocity addition law.  Comments on "Deriving relativistic momentum and energy: II,  Three-dimensional case_.  European Journal of Physics, 27:L17-L20


# Further information
For more detail, see the package vignette

```
vignette("lorentz")
``` 
 

