---
title: "mobius"
author: "Robin KS Hankin"
date: "11/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Möbius addition 

We define Möbius addition $\oplus$ on the unit disk $\mathbb{D} =
\left\lbrace z\in\mathbb{C}\colon\left|z\right|<1\right\rbrace$ as
follows:

\begin{equation}
a\oplus z = \frac{a+z}{1+\overline{a}z}
\end{equation}

(naturally we define $a\ominus z=a\oplus(-z)$ and it convenient to
write $\ominus z=0\ominus z=-z$).



The $\oplus$ operator is neither commutative nor associative ($a\oplus
b\neq b\oplus a$ and $a\oplus(b\oplus c)\neq (a\oplus b)\oplus c$ in
general).  However, it does posess the _automorphic inverse property_

\begin{equation}
\ominus(a\oplus b) = \ominus a\ominus b
\end{equation}

and the _left cancellation law_

\begin{equation}
\ominus a\oplus (a\oplus z) = z
\end{equation}


proof:
\[
\ominus(a\oplus b) = -(a\oplus b)=-\frac{a+b}{1+\overline{a}b}=
=\frac{(-a)+(-b)}{1+\overline{a}b}
=\frac{(-a)+(-b)}{1+\overline{(-a)}(-b)}
=(-a)\oplus(-b)
=\ominus a\ominus b
\]

\[
\ominus a\oplus (a\oplus z) = 
\ominus a\oplus\frac{a+z}{1+\overline{a}z}
=\frac{-a + \frac{a+z}{1+\overline{a}z}}{1+\overline{(-a)}\frac{a+z}{1+\overline{a}z}}
=\frac{-a\overline{a}z +z}{1-\overline{a}a}
=\frac{z(1-a\overline{a})}{1-\overline{a}a}
=z,
\]
the last equality following from the fact that $a\in\mathbb{D}$.


```{r}

`%+%` <- function(x,y){(x+y)/(1+Conj(x)*y)}
`%-%` <- function(x,y){(x-y)/(1-Conj(x)*y)}

square <- function(a,b,len=100){ # a is lower left point, b is upper right
    a <- a+0i
    b <- b+0i
    c(
    seq(from=Re(a) + 1i*Im(a), to=Re(b) + 1i*Im(a),len=len),
    seq(from=Re(b) + 1i*Im(a), to=Re(b) + 1i*Im(b),len=len),
    seq(from=Re(b) + 1i*Im(b), to=Re(a) + 1i*Im(b),len=len),
    seq(from=Re(a) + 1i*Im(b), to=Re(a) + 1i*Im(a),len=len)
    )
}

gyrosetup <- function(...){
	par(pty="s")
	plot(NA,xlim=c(-1,1),ylim=c(-1,1),asp=1,axes=FALSE,xlab="",ylab="")
theta <- seq(from=0,to=2*pi,len=400)
points(sin(theta),cos(theta),type="l",col="gray",...)
segments(x0=-1, x1= +1,y0=0,col='gray')
segments(x0=0,y0=-1,y1=+1,col='gray')
text(0.9,0.06,"Re")
text(0.1,0.9,"Im")
}

polyz <- function(z,...){polygon(x=Re(z),y=Im(z),...)}

gyrosetup()

S <- square(-1-1i,1+1i)/6
points(S,type="l")

zs <- c(
    +0.2 + 0.38i,
    -0.2 + 0.58i,
    +0.2 - 0.38i,
    -0.3 - 0.38i
)*1.4

for(z in zs){
points(Re(z),Im(z))
polyz(S %+% z,col=rgb(1,0,0,alpha=0.3))
polyz(z %+% S,col=rgb(0,0,1,alpha=0.3))
}

```
